---
title: 'INFO 550: Project Markdown'
author: "Ian Fowler"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(dplyr)
library(labelled)
library(sjlabelled)
library(gtsummary)
library(ggplot2)
library(parameters)
library(kableExtra)
```

```{r, echo = FALSE}
here::i_am('final_project.Rmd')
```

```{r, echo = FALSE}
project.data4 <- readRDS(
  here::here("output/project_data4.rds")
)

table1 <- readRDS(
  here::here("output/table1.rds")
)

fig1 <- readRDS(
  here::here("output/fig1.png")
)
```

## Introduction

In 2009 a variant of the seasonal flu spread globally beginning a pandemic. H1N1, also known as "Swine Flu" due to the source of transmission, spread to every region and caused approximately __94481__ cases and __429__ deaths from April 2009 to July 2010 alone.

In this analysis we will investigate the relationships between __geographic region__, __total number of cases__, and __total number of deaths__ associated with the H1N1 pandemic. While the entire pandemic lasted from January 2009 to August of 2010, the data set sourced from [Kaggle](https://www.kaggle.com/datasets/imdevskp/h1n1-swine-flu-2009-pandemic-dataset?resource=download) as provided by the WHO only contains data from April 2009 to July 2010 as after July, countries were no longer required to submit individual level data and the date for after July is not complete.

## Data Aggregation

The data from the data set listed at the country level with daily cumulative case and death counts, For the analysis, the country level data was aggregated to the regional level with 8 defined regions (**North America**, **Central America & the Caribbean**, **South America**, **Europe**, **Africa**, **Middle East**, **Asia**, and **Oceania**) in order to compare case and death counts between regions (using nominal variables). 

The total case and death counts were also calculated for each of the regions for further analysis.

## Descriptive Analysis

The data calculated by the regional aggregation is described below in the following table and plot. The total number of cases and deaths per region is shown and the correlation between number of cases and number of deaths is illustrated.

```{r}
table1
```

<center>
*(1)North America (2)Central America & the Caribbean (3)South America (4)Europe (5)Africa (6)Middle East (7)Asia (8)Oceania*
<center>

```{r}
#| fig.align = "center",
#| out.width = "700px"
fig1
```

The plot shows the __total number of cases__ for each region along the x-axis with the __total number of deaths__ for each region described by the color gradient.
